---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# celavi

<!-- badges: start -->
[![R-CMD-check](https://github.com/jbkunst/celavi/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jbkunst/celavi/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of celavi is to join the main features of two functions  
that I use _really_ often `vip::vi_permute` and `DALEX::model_parts`. Both functions
do the _same_ task of calculate _drop out loss via permutation_, 
but they have different features and approach.

In the case of `vip::vi_permute` is more direct to use (imho), have an implementation for
parallel processing, can be used with a `sample_frac` parameter. Otherwise, in the case of `DALEX::model_parts` I like the user can give custom `metric`s as a loss functions,
the _base line_ and _full model_ references values, and the plots. 

To that features I added some features to my _personal_ taste.

- Add progress bars to the sequential and parallel process using `progress::progress_bar` and `progressr::progress`
- Give the possibility of to the user to access to the _raw_ data.
- Verbose information using `cli::cli_alert_info`.

## References

The `vip` package from [koalaverse](https://github.com/koalaverse), and the 
`DALEX` package from [MI²](https://www.mi2.ai/). In particular these links are awesome: https://koalaverse.github.io/vip/articles/vip.html
and https://ema.drwhy.ai/featureImportance.html#featureImportanceR.

Please, visit the links and used that awesome tools!

## Installation

You can install the development version of celavi from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jbkunst/celavi")
```

## Example

```{r example}
library(celavi)

lm_model <- lm(mpg ~ ., data = mtcars)

set.seed(123)

vi <- celavi::variable_importance(lm_model, data = mtcars, iterations = 10)

dplyr::glimpse(vi)

nrow(vi)
# nrow(vi) = (ncol(mtcars) - 1 + 2) * iterations

plot(vi)
```
