---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning = FALSE
)
```

# celavi

`celavi` provides permutation-based variable importance and feature selection tools.

The package is inspired by two functions I use often: `vip::vi_permute()` and
`DALEX::model_parts()`. Both estimate variable importance by measuring the
drop-out loss after permuting predictors, but they have different strengths.

`vip::vi_permute()` is direct to use, supports parallel processing, and works
well with sampling arguments such as `sample_frac`. `DALEX::model_parts()` makes
it easy to use custom loss functions, compare against baseline and full-model
reference values, and produce useful plots.

`celavi` keeps the parts I like from both approaches and adds a few conveniences:

- progress bars for sequential and parallel runs;
- access to the raw permutation results;
- informative messages through `cli`;
- a lightweight iterative feature-selection workflow.

## References

`celavi` is inspired by:

- `vip`, from the koalaverse: https://koalaverse.github.io/vip/articles/vip.html
- `DALEX`, from MI²: https://ema.drwhy.ai/featureImportance.html#featureImportanceR

Both are excellent packages. `celavi` is not intended to replace them; it is a
small package focused on my preferred workflow for permutation importance and
feature selection.

## Installation

You can install the package from GitHub with:

```r
pak::pkg_install("jbkunst/celavi")
```

## Example I: Variable Importance

```{r example}
library(celavi)

lm_model <- lm(mpg ~ ., data = mtcars)

set.seed(123)

vi <- celavi::variable_importance(lm_model, data = mtcars, iterations = 100)

dplyr::glimpse(vi)

nrow(vi)
# nrow(vi) == (ncol(mtcars) - 1 + 2) * iterations

plot(vi)
```

And compare with other model.

```{r}
rf <- randomForest::randomForest(mpg ~ ., data = mtcars)

vi_rf <- celavi::variable_importance(rf, data = mtcars, iterations = 100)

plot(vi, vi_rf)
```

The previous chart shows that the random forest has a smaller, better RMSE.
It is also less affected by permuting some predictors. For example, permuting
`wt` has a visible impact on the linear model.

## Example II: Feature Selection

```{r}
set.seed(123)

data(credit_data, package = "modeldata")

credit_data <- credit_data[complete.cases(credit_data),]
credit_data$Status <- as.numeric(credit_data$Status) - 1

# convert factor to dummies (to compare results with glmnet)
credit_data <- as.data.frame(model.matrix(~ . - 1, data = credit_data))

trn_tst <- sample(
  c(TRUE, FALSE),
  size = nrow(credit_data),
  replace = TRUE, 
  prob = c(.7, .3)
  )

credit_data_trn <- credit_data[ trn_tst,]
credit_data_tst <- credit_data[!trn_tst,]

fs <- feature_selection(
  glm,
  credit_data_trn,
  response = "Status",
  stat = min,
  iterations = 10,
  sample_frac = 1, 
  predict_function = predict.glm,
  # function accepts specific argument for the fit function
  family  = binomial
)

fs

plot(fs)
```

The result is a simpler model with little apparent loss in predictive performance.

Now we can compare with some other feature selection techniques.

```{r, results='hide'}
mod_fs <- attr(fs, "final_fit")

mod_full <- glm(Status ~ ., data = credit_data_trn, family = binomial)

mod_step <- step(mod_full, trace = FALSE) 

# wrapper around glmnet::cv.glmnet()
mod_lasso <- risk3r::featsel_glmnet(mod_full, plot = FALSE)
```

```{r}
models <- list(
  "featsel by vip" = mod_fs,
  "stepwise"  = mod_step,
  "lasso"     = mod_lasso
)

dmetrics <- purrr::map_df(
  models,
  risk3r::model_metrics,
  newdata = credit_data_tst, 
  .id = "method"
)

dmetrics
```

Not the best model in terms of metrics. But if we see the number of coefficients:

```{r}
dnvars <- purrr::map_df(
  models,
  ~ tibble::tibble(`# variables` =  length(coef(.x))),
  .id = "method"
)

dplyr::full_join(dnvars, dmetrics, by = dplyr::join_by(method))
```

We can check the loss in each iteration, so you can choose what combinations
of loss/number of variables you want.

```{r, fig.height=14}
do.call(plot, attr(fs, "variable_importance")) +
  ggplot2::scale_y_continuous(
    breaks = scales::pretty_breaks(7),
    sec.axis = ggplot2::dup_axis(~ 1 - .x, name = "AUC", labels = scales::percent)
  )

fs
```



