---
title: "Feature selection on MNIST data"
editor_options:
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 5
)
```

```{r setup}
library(celavi)
library(dplyr)
library(ggplot2)
library(klassets)
library(ranger)
library(tidyr)
```

# Feature selection on MNIST data

This article shows how `celavi` can be used to visualize the trade-off between
predictive performance and model simplicity.

The example uses a reduced version of the MNIST data. The goal is not to build
the best possible MNIST classifier, but to show how the feature-selection
procedure removes variables across rounds and how much predictive performance is
lost in the process.

```{r parameters}
set.seed(123)

n_train  <- 10000
n_test   <- 1000
n_pixels <- 100
iter     <- 15
frac     <- 1

num_trees     <- 100
max_depth     <- 8
min_node_size <- 20

stat_fun <- median
```

The experiment uses `r n_train` training observations, `r n_test` test
observations, and `r n_pixels` randomly selected pixel variables.

The argument `iter = r iter` controls how many permutation losses are estimated
for each variable in each selection round. It does not fix the number of
selection rounds. The number of rounds depends on how many variables are removed
after each step.

```{r data}
pixel_cols <- sample(setdiff(names(klassets::mnist_train), "label"), n_pixels)

data <- klassets::mnist_train |>
  dplyr::mutate(label = factor(label)) |>
  dplyr::slice_sample(n = n_train) |>
  dplyr::select(label, dplyr::all_of(pixel_cols))

test <- klassets::mnist_test |>
  dplyr::mutate(label = factor(label)) |>
  dplyr::slice_sample(n = n_test) |>
  dplyr::select(label, dplyr::all_of(pixel_cols))
```

# Feature selection procedure

At each round, `celavi` fits a model, estimates the loss caused by permuting each
variable, and removes variables that do not contribute enough to predictive
performance.

Because `label` is a multiclass factor, the default loss function is based on
classification accuracy. Lower loss values indicate better predictive
performance.

```{r feature-selection}
x <- feature_selection(
  ranger::ranger,
  data = data,
  test = test,
  response = "label",
  stat = stat_fun,
  iterations = iter,
  sample_frac = frac,
  predict_function = function(object, newdata) {
    predict(object, data = newdata)$predictions
  },
  parallel = FALSE,
  # ranger-specific arguments
  num.trees = num_trees,
  max.depth = max_depth,
  min.node.size = min_node_size,
  num.threads = 1
)
x
```

```{r summary}
selection_summary <- x |>
  dplyr::transmute(
    round,
    loss = mean_value,
    accuracy = 1 - mean_value,
    n_variables
  )

selection_summary
```

# Selection path

```{r selection-path}
loss_data <- x |>
  dplyr::select(round, values) |>
  tidyr::unnest(values) |>
  dplyr::transmute(
    round,
    metric = "Loss function",
    value = values
  )

variables_data <- x |>
  dplyr::transmute(
    round,
    metric = "Number of variables",
    value = n_variables
  )

plot_data <- dplyr::bind_rows(loss_data, variables_data)

ggplot2::ggplot() +
  ggplot2::geom_boxplot(
    data = dplyr::filter(plot_data, metric == "Loss function"),
    ggplot2::aes(round, value, group = round)
  ) +
  ggplot2::geom_col(
    data = dplyr::filter(plot_data, metric == "Number of variables"),
    ggplot2::aes(round, value)
  ) +
  ggplot2::facet_wrap(ggplot2::vars(metric), scales = "free_y") +
  ggplot2::labs(
    x = "Selection round",
    y = NULL
  ) +
  ggplot2::theme_minimal()
```

The left panel shows the distribution of the loss function across selection
rounds. The right panel shows the number of variables kept by the model.

The expected trade-off is visible: the number of predictors falls sharply, while
the loss function changes as the model is refitted with fewer variables.

The loss does not need to increase monotonically. After each round, the model is
trained again using the remaining variables, and removing weak or noisy pixels
can sometimes keep the loss stable. The main purpose of the example is to show
how `celavi` makes this selection path visible.

```{r final-summary}
first_round <- selection_summary |>
  dplyr::slice_min(round, n = 1)

last_round <- selection_summary |>
  dplyr::slice_max(round, n = 1)

initial_variables <- first_round$n_variables
final_variables <- last_round$n_variables

initial_accuracy <- first_round$accuracy
final_accuracy <- last_round$accuracy

variable_reduction <- 1 - final_variables / initial_variables
accuracy_change_pp <- 100 * (final_accuracy - initial_accuracy)
```

In this run, the model starts with `r initial_variables` variables and finishes
with `r final_variables` variables. This is a reduction of approximately
`r scales::percent(variable_reduction, accuracy = 1)` of the original predictors.

The estimated predictive accuracy changes from approximately
`r scales::percent(initial_accuracy, accuracy = 0.1)` to
`r scales::percent(final_accuracy, accuracy = 0.1)`. This is a change of about
`r round(accuracy_change_pp, 1)` percentage points.

This article is intentionally small enough to render quickly. The objective is
not to optimize MNIST performance, but to make the feature-selection path easy to
inspect.